---
title: "Figure 1B: Staphylococcal species by body-sites PCoA plot using Bray-Curtis"
author: "Payal Joglekar"
date: "`r Sys.Date()`"
output: html_document
---

===============================================================================

###Background:
Take the phyloseq object *ps_staph_ASV_filter_final.rds* created by subsetting ps_HV_site_time_final.ps.rds phyloseq object on Genus = Staphylococcus. 
The purpose of this of this document is to create Figure1C showing a PCoA plot of ASVs driving body-site separation based on bray-curtis <br />

###Steps:
#### 1.1 Load libraries and set parameters
````{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

```{r, results='hide'}
library(phyloseq); packageVersion("phyloseq")
library(tidyverse)
library(gridExtra)
library(DESeq2)
library(dplyr)
library(vegan)
library(RColorBrewer)
library(ggplot2)
library(cluster)
library(ggdendro)
library(grid)
library(ggrepel)
library(data.table)
library(reshape2)
theme_set(theme_bw())
```

```{r, include=FALSE}
path <- getwd()
```

### 1.2 Importing phyloseq object 
```{r}
ps_staph <- readRDS(file.path(path,"ps_staph_ASV_filter_final.rds"))
ps_staph #check the phyloseq object

```

####1.3  Define a mapping file from staph phyloseq sample_data
```{r}
map_staph <- data.frame(sample_data(ps_staph))
map_staph$Subject_ID<-as.numeric(levels(map_staph$Subject_ID))[map_staph$Subject_ID]
```

####1.4 Define color scheme
```{r}
species_colors_3 <- c("S. epidermidis" = "#1D2951", "S. capitis" = "#31afde", "S. hominis" = "#b1e1ae", "S. warneri" = "#ff9999", "S. haemolyticus"= "#056f00", "S. lugdunensis" = "#9e9d24", "S. saccharolyticus" = "#628df2" , "S. cohnii" = "#febf00", "S. auricularis" = "#ff8400", "S. saccharolyticus group" = "#6146ca", "S. aureus" = "#d30000", "S. pettenkoferi" = "#c35817", "S. petrasii" = "#c89369", "S. haemolyticus group" = "#60a130", "S. pasteuri"= "#de6f6f",  "S. saprophyticus" = "#fde403", "S. simulans" = "#592c4b") 

sites_colors_6 <- c("N" = "#98CBAC", "Ac" = "#DFEFBF", "Pc" = "#C5E28C", "Gc" = "#ACC973", "Ax" = "#EBEBC0", "Ic" = "#DBDB8D", "Um" = "#969659", "Ph" = "#FEE5BB", "Tw" = "#FDCF80", "Tn" = "#887960", "Hp" = "#EDC0C0", "Vf" = "#DE8C8C", "Ba" = "#CEF3EF", "Mb" = "#93E4DB", "Oc" = "#D9E9F5", "Gb" = "#A6CCE9", "Al" = "#73A3C9",  "Ra" = "#53636F", "Ea" = "#CABFEF")

#Define plot features
add_barplot_elements <- function(x) {
      theme_update() +
        theme(plot.title = element_text(color="black", size=12, face="bold.italic"),
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        axis.text.x = element_text(size= 10, angle=90, hjust=1),
        axis.text.y = element_text(size = 10),
        strip.text.x = element_text(color="black", size=10, face="bold"),
        panel.spacing = unit(0.25, "lines"),
        strip.background = element_rect(colour = "black", fill = "grey90"),
        strip.placement = "outside",
        legend.title = element_text(color = "black", size = 12),
        legend.text = element_text(color = "black", size = 10)) 
}

```


#### 2. Data Transformation

Data tranformation to achieve variance stabilization
```{r}
library(vsn)
##Variance vs mean of raw data
otus.raw <- data.frame(otu_table(ps_staph))
raw = meanSdPlot(as.matrix(otus.raw), rank=FALSE) 

##Transformation##
library(vegan)
#Simple proportion
psRA <- transform_sample_counts(ps_staph, function(x) x/sum(x))

#arcsine tranform
psAST <- transform_sample_counts(psRA, function(x) {asin(sqrt(x))})

#Set up an otu tables with arcsine transformations
otus.AST = data.frame(otu_table(psAST))

#plot arcsine tranform
AST= meanSdPlot(as.matrix(m), rank=FALSE) 

```


#### 3. PCoA: bray-curtis
```{r}

m <- data.frame((otu_table(psAST)))

d.all <- vegdist(m, method='bray', binary=F) #Bray-Curtis 

pcoa_all <- cmdscale(d.all, k=4, eig = TRUE)
pcoa <- data.frame(pcoa_all$points, stringsAsFactors=F)

pcoa1 <- data.frame(t(pcoa))
pcoa <- data.frame(t(pcoa1))

eig <- pcoa_all$eig
loadings <- (eig/sum(eig))[1:4]
colnames(pcoa) <- c('axis1', 'axis2', 'axis3', 'axis4')

# pcoa axis 1,2,3 loading 
loadings 
loadings <- as.data.frame(loadings)
loadings[1,]
loadings[2,]
loadings[3,]
loadings[4,]
```

#### 4. Covariant analysis
```{r}
df.1<-data.frame(m, stringsAsFactors=F) 
df.2 <- data.frame(t(df.1), stringsAsFactors=F) #transpose 

## This step will give you the statistically significant taxa (genera) 
fit_taxa <- envfit(ord=pcoa, df.1, permutations=999) #p.max  
fit_taxa

fitted <- data.frame(fit_taxa$vectors$arrows[p.adjust(fit_taxa$vectors$pvals, method='fdr')<0.01,]*sqrt(fit_taxa$vectors$r[p.adjust(fit_taxa$vectors$pvals, method='fdr')<0.01]), stringsAsFactors=F) #adjusted p-value<0.05

fitted_ASV <- data.frame(rownames(fitted))

fitted <- fitted %>% mutate(taxa = fitted_ASV$rownames.fitted.)
```


#### 5. Metadata
```{r}
## loading metadata 
metadata <- data.frame(sample_data(psAST),stringsAsFactors=F)

#Transfer information of the factors from metadata to pcoa table  
rownames(pcoa) <- rownames(metadata)

pcoa <- tibble::rownames_to_column(pcoa, 'Sample')
metadata <- tibble::rownames_to_column(metadata, 'Sample')


pcoa <- full_join(metadata, pcoa, by = 'Sample')
```

#### 6. Plot

```{r}

ggplot(pcoa) + 
  geom_point(aes(x=axis1, y=axis2,fill=Site),colour="black",pch=21,size=4, alpha = 3) +
  geom_segment(data=fitted, aes(x=0, xend=axis1/2, y=0, yend=axis2/2), 
               arrow=arrow(length=unit(0.15, 'cm')), color='black', alpha=1) +
  #geom_label_repel(data=fitted, aes(x=axis1/2, y=axis2/2, label=taxa), alpha=0.8, size=3, color='black', direction='both', nudge_y=0.2, segment.size=0.1, xlim=c(-0.75, 0.75)) +
  ggtitle("") +  #PCoA with Bray-Curtis
  scale_fill_manual(values = sites_colors_6) +
  xlab(paste("PC1 (", round(100*loadings[1,1],2), "%)", sep='')) + 
  ylab(paste("PC2 (", round(100*loadings[2,1],2), "%)", sep='')) +
  xlim(-0.6,0.6) + ylim(-0.5,0.5) + 
  theme_bw() + 
  theme(legend.text=element_text(size=20), 
        plot.title = element_text(face="bold",size=20,hjust=0.5)) + 
  guides(fill=guide_legend(title=NULL))

```


#### 7. Statistics
```{r}
#PERMANOVA

set.seed(1)

# Calculate the desired distance matrix
d <- phyloseq::distance(psAST, method = "bray")

# make a data frame from the sample_data
sampledf <- data.frame(sample_data(ps_staph))

# Adonis test
ad <- adonis(d ~ Site, sampledf, permutations = 10000)
ad

```
