---
title: "Figure 2A: Staphylococcus genus pangenome core gene calculation"
author: "Payal Joglekar"
date: "`r Sys.Date()`"
output: html_document
---

Generate files for calculating genes (core + accessory) in all sub-sets of the genus pan-genome (for Figure 2B)

Steps:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

1. Load libraries

```{r cars}
library(tidyverse)
library(RColorBrewer)
library(reshape2)
library(data.table)
```

```{r, include=FALSE}
path <- getwd()
```


2. Upload files and tidy them

```{r}
#pananroo gene_presence_absence
df <- fread(file.path(path,"genus_gene_presence_absence.csv"), data.table = FALSE) %>% as_tibble()

df1 <- column_to_rownames(df, var = "ID")
df1 <- df1 %>% select(3:128)

#metadata
meta <- read.csv(file.path(path,"metadata.csv"), header=T)
 
#trim metadata
meta <- meta %>% dplyr::rename(subject_ID = "subject.number.for.manuscript")
meta <- meta %>% select(c(isolate, genome_trim, species, subject_ID, site))

#change genome_trim names to isolate names
df1t <- as.data.frame(t(df1))
df1t <- rownames_to_column(df1t, var = "genome_trim")
meta1 <- meta %>% select(isolate, genome_trim)
df1t1 <- left_join(meta1, df1t, by = 'genome_trim')
df1t1 <- df1t1 %>% select(!genome_trim)
df1t1 <- column_to_rownames(df1t1, var = "isolate")
```

3.1 Genus pan-genome categories (Figure 2A)
```{r}
#create a new column with # of genomes carrying a gene
df2 <- as.data.frame(t(df1t1)) #transpose
df2 <- df2 %>% mutate(sum = rowSums(df2)) #this gives number of genomes, temporary column
df2$sum <- as.numeric(df2$sum)

#total genes
nrow(df2)
#core
c <- nrow(df2 %>% filter(sum >= 126))
#soft-core
sc <- nrow(df2 %>% filter(sum >= 120 & sum < 126))
#shell
s <- nrow(df2 %>% filter(sum >= 19 & sum < 120))
#cloud
r <- nrow(df2 %>% filter(sum < 19))

c(c, sc, s, r)
```

3.2 Subset df onto each species (for Figure 2B)
```{r}
#all
dfall <- df2 
dfall <- dfall %>% filter(sum > 0)
dfall <- rownames_to_column(dfall, var = "gene")

#capitis
metac <- meta %>% filter(species == "S. capitis")
dfc <- df2 %>% select(metac$isolate)
dfc <- dfc %>% mutate(sum = rowSums(dfc)) 
dfc <- dfc %>% filter(sum > 0)
dfc <- rownames_to_column(dfc, var = "gene")

#epidermidis
metae <- meta %>% filter(species == "S. epidermidis")
dfe <- df2 %>% select(metae$isolate)
dfe <- dfe %>% mutate(sum = rowSums(dfe)) 
dfe <- dfe %>% filter(sum > 0)
dfe <- rownames_to_column(dfe, var = "gene")

#hominis
metah <- meta %>% filter(species == "S. hominis")
dfh <- df2 %>% select(metah$isolate)
dfh <- dfh %>% mutate(sum = rowSums(dfh)) 
dfh <- dfh %>% filter(sum > 0)
dfh <- rownames_to_column(dfh, var = "gene")
```

3.3 Genes restricted to 2 species (core + accessory) (for Figure 2B)
```{r}
#cap+epi all
ce_all <- inner_join(dfc, dfe, by = 'gene') 
ce_all1 <- anti_join(ce_all, dfh, by = 'gene')
ce_all_df <-df %>% filter(ID %in% ce_all1$gene) 
write_csv(ce_all_df, "genus_pangenome_scap_sepi_all.csv")

#cap+epi core (species-restricted-core)
ce_core <- ce_all1 %>% filter(sum.x >= 21 & sum.y >= 47)
ce_core_df <- df %>% filter(ID %in% ce_core$gene) 

##############################
#cap+hom all
ch_all <- inner_join(dfc, dfh, by = 'gene') 
ch_all1 <- anti_join(ch_all, dfe, by = 'gene')
ch_all_df <-df %>% filter(ID %in% ch_all1$gene)
write_csv(ch_all_df, "genus_pangenome_scap_shom_all.csv")

#cap+hom core (species-restricted-core)
ch_core <- ch_all1 %>% filter(sum.x >= 21 & sum.y >= 53)
ch_core_df <- df %>% filter(ID %in% ch_core$gene) 

##############################
#epi+hom all
eh_all <- inner_join(dfe, dfh, by = 'gene') 
eh_all1 <- anti_join(eh_all, dfc, by = 'gene')
eh_all_df <-df %>% filter(ID %in% eh_all1$gene)
write_csv(eh_all_df, "genus_pangenome_sepi_shom_all.csv")

#epi+hom core (species-restricted-core)
eh_core <- eh_all1 %>% filter(sum.x >= 47 & sum.y >= 53)
eh_core_df <- df %>% filter(ID %in% eh_core$gene) 
```


3.4 Genes restricted to 1 species (core + accessory) (for Figure 2B)
```{r}
#cap
cap_all <- anti_join(dfc, dfe, by = 'gene')
cap_all1 <- anti_join(cap_all, dfh, by = 'gene')
cap_all_df <-df %>% filter(ID %in% cap_all1$gene)
write_csv(cap_all_df, "genus_pangenome_scap_all.csv")

#core (species-restricted-core)
cap_core <- cap_all1 %>% filter(sum >= 21)
cap_core_df <- df %>% filter(ID %in% cap_core$gene) 

##############################
#epi
epi_all <- anti_join(dfe, dfc, by = 'gene')
epi_all1 <- anti_join(epi_all, dfh, by = 'gene')
epi_all_df <-df %>% filter(ID %in% epi_all1$gene)
write_csv(epi_all_df, "genus_pangenome_sepi_all.csv")

#core (species-restricted-core)
epi_core <- epi_all1 %>% filter(sum >= 47)
epi_core_df <- df %>% filter(ID %in% epi_core$gene) 

##############################
#hom
hom_all <- anti_join(dfh, dfc, by = 'gene')
hom_all1 <- anti_join(hom_all, dfe, by = 'gene')
hom_all_df <-df %>% filter(ID %in% hom_all1$gene)
write_csv(hom_all_df, "genus_pangenome_shom_all.csv")

#core (species-restricted-core)
hom_core <- hom_all1 %>% filter(sum >= 53)
hom_core_df <- df %>% filter(ID %in% hom_core$gene)
##############################
```

3.5 Genes in 3 species (core plus accessory) (for Figure 2B)
```{r}
#core to all three species (genus-core)
staph_all_core <- dfall %>% filter(sum >= 120)
staph_all_core_df <- df %>% filter(ID %in% staph_all_core$gene) 
write_csv(staph_all_core_df, "genus_pangenome_staph_genus_core.csv")


#genus_shell
staph_all_shell <- dfall %>% filter(sum >= 19 & sum < 120)
staph_all_shell_df <- df %>% filter(ID %in% staph_all_shell$gene) 
write_csv(staph_all_shell_df, "genus_pangenome_staph_genus_shell.csv")

#genus-cloud
staph_all_cloud <- dfall %>% filter(sum < 19)
staph_all_cloud_df <- df %>% filter(ID %in% staph_all_cloud$gene) 
write_csv(staph_all_cloud_df, "genus_pangenome_staph_genus_cloud.csv")

#accessory to all three species
non_all_core_list <- as.vector(c(ce_all1$gene, ch_all1$gene, eh_all1$gene, cap_all1$gene, epi_all1$gene, hom_all1$gene, staph_all_core$gene))
staph_all_accessory <- dfall %>% filter(!gene %in% non_all_core_list)
staph_all_accessory_df <- df %>% filter(ID %in% staph_all_accessory$gene)
write_csv(staph_all_accessory_df, "genus_pangenome_staph_genus_accessory.csv")

#all accessory genes (except 1660 core)
staph_non_core <- dfall %>% filter(!gene %in% staph_all_core$gene) 
staph_non_core_df <- df %>% filter(ID %in% staph_non_core$gene) 
write_csv(staph_non_core_df, "genus_pangenome_staph_all_non_core.csv")


#genes common to all irrespective of core or accessory
non_all_list <- as.vector(c(ce_all1$gene, ch_all1$gene, eh_all1$gene, cap_all1$gene, epi_all1$gene, hom_all1$gene))
staph_all <- dfall %>% filter(!gene %in% non_all_list)
staph_all_df <- df %>% filter(ID %in% staph_all$gene)
write_csv(staph_all_df, "genus_pangenome_staph_common_to_all_three.csv")

```
