---
title: "Figure S3: Mean relative abundance boxplots for core staphylococcal species"
author: "Payal Joglekar"
date: "`r Sys.Date()`"
output: html_document
---

###Background:
Take the phyloseq object *ps_staph_ASV_filter_final.rds* created by subsetting ps_HV_site_time_final.ps.rds phyloseq object on Genus = Staphylococcus.  
<br />
The purpose of this of this document is to create a mean relative abundance barplot of the three major Staphylococcus species (Figure S3)


###Steps:
#### 1.1 Load libraries and set parameters

````{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

### 1.1 Load libraries 
```{r, results='hide'}
library(phyloseq); packageVersion("phyloseq")
library(Biostrings); packageVersion("Biostrings")
library(tidyverse)
library("RColorBrewer") 
theme_set(theme_bw())
```

```{r, include=FALSE}
path <- path <- getwd()
```

####1.2 Importing phyloseq object 
```{r}
ps_staph <- readRDS(file.path(path,"ps_staph_ASV_filter_final.rds"))
ps_staph #check the phyloseq object
```


####1.3  Define a mapping file from staph phyloseq sample_data
```{r}
map_staph <- data.frame(sample_data(ps_staph))
map_staph$Subject_ID<-as.numeric(levels(map_staph$Subject_ID))[map_staph$Subject_ID]
```

####1.4 Define color scheme
```{r}
species_colors_3 <- c("S. epidermidis" = "#1D2951", "S. capitis" = "#31afde", "S. hominis" = "#b1e1ae") 

sites_colors_9 <- c("N" = "#98CBAC", "Ac" = "#DFEFBF", "Pc" = "#C5E28C", "Gc" = "#ACC973", "Ax" = "#EBEBC0", "Ic" = "#DBDB8D", "Um" = "#969659", "Ph" = "#FEE5BB", "Tw" = "#FDCF80", "Tn" = "#887960", "Hp" = "#EDC0C0", "Vf" = "#DE8C8C", "Ba" = "#CEF3EF", "Mb" = "#93E4DB", "Oc" = "#D9E9F5", "Gb" = "#A6CCE9", "Al" = "#73A3C9",  "Ra" = "#53636F", "Ea" = "#CABFEF")

#Define plot features
add_barplot_elements <- function(x) {
      theme_update() +
        theme(plot.title = element_text(color="black", size=12, face="bold.italic"),
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        axis.text.x = element_text(size= 10, angle=90, hjust=1),
        axis.text.y = element_text(size = 10),
        strip.text.x = element_text(color="black", size=10, face="bold"),
        panel.spacing = unit(0.25, "lines"),
        strip.background = element_rect(colour = "black", fill = "grey90"),
        strip.placement = "outside",
        legend.title = element_text(color = "black", size = 12),
        legend.text = element_text(color = "black", size = 10)) 
}

```


####2. Merge all ASV from the same species using the tax_glom function
```{r}
psglom <- tax_glom(ps_staph, taxrank="Species", NArm = TRUE)
psglom

map_psglom <- data.frame(sample_data(psglom)) #make a metadata dataframe
map_psglom$SeqDepths <- sample_sums(psglom)
```


####4 We wil now generate boxplot of select species across different body sites
```{r}
# Transform Taxa counts to relative abundance
ra <- transform_sample_counts(psglom, function(OTU) OTU/sum(OTU) * 100)
taxa_tbl <- psmelt(ra)

#species to subset
taxa_keep <- c("Staphylococcus_hominis", "Staphylococcus_&capitis&caprae", "Staphylococcus_epidermidis") #option one

#trim by species and/or site
taxa_tbl_spp <- taxa_tbl %>% filter(Species %in% taxa_keep)


#site order
taxa_tbl_spp$Site <- factor(taxa_tbl_spp$Site, levels = c("Ea", "Oc", "Gb", "Ra", "Ba", "Mb", "Hp", "Vf",  "N", "Ac", "Ic", "Tw", "Ph" , "Tn"))

#rename species
taxa_tbl_spp$SppRename[taxa_tbl_spp$Species == 'Staphylococcus_hominis'] <- 'S. hominis'
taxa_tbl_spp$SppRename[taxa_tbl_spp$Species == 'Staphylococcus_&capitis&caprae'] <- 'S. capitis'
taxa_tbl_spp$SppRename[taxa_tbl_spp$Species == 'Staphylococcus_epidermidis'] <- 'S. epidermidis'


#combined species boxplot
bp_species <- taxa_tbl_spp %>% 
  ggplot(aes(x = Site, y = Abundance, fill = SppRename)) +
  geom_boxplot() +
  #facet_wrap(~ SiteType, scales = "fixed", nrow = 1) +
  scale_fill_manual(values = species_colors_3) +
  labs(x = "Body site",
       y = "Mean Relative Abundance",
       title = "Species Relative Abundance")  +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 14),
    legend.position = "bottom"
  ) 

bp_species
```


